Oriented Gradient Histogram applied to P300 Detection

### 2017-12-04 response.letter.tex

Another issue with the manuscript is its mathematical vagueness. For example, the variable T is used as a vector or two-tuple but its notation is different from the notation the authors use for vectors. Moreover, there is something wrong in equation (4) and something inconsistent between equations (7) and (8). Also, the authors go back and forth the digital space and the continuous space and although there is an equivalence, the manuscript could benefit from a clearer explanation.

Another issue that I find worrisome is the design of experiments. First of all, it seems that the number of subjects is rather low, it would be interesting to know why the authors decided to use such a subset. Also, the authors speculate that their results are not that good because they did not take into account the latency and magnitude of the signals, there is previous evidence in the literature that the P300 suffers from these, and other, problems and, thus, one wonders why the authors did not anticipate such problems.

A couple of finale observations. First, it seems the authors assume that the readers have some previous deep knowledge of some topics such as that some channels should carry more or better information without a previous reference. Another issue has to do with the way the manuscript is written, sometimes is very difficult to follow the ideas the authors want to convey, an example can be found in the paragraph containing equation (5), the manuscript surely benefit from a careful review before submitting.

\subsection*{\ovalbox{General Comments}}
I have read the manuscript "EEG characterization and classification based on Histogram of Gradients of Signal Plots" with great interest since its title promised a solution to an important problem. In spite of presenting an interesting methodology and some interesting issues related to EEG characterization, I consider the manuscript is not ready for publication. 
\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for your feedback. We have modified several points according to your recommendations, we hope the article is now better suited for publication. 
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

### 2017-12-04 response.letter.tex

A couple of finale observations. First, it seems the authors assume that the readers have some previous deep knowledge of some topics such as that some channels should carry more or better information without a previous reference. Another issue has to do with the way the manuscript is written, sometimes is very difficult to follow the ideas the authors want to convey, an example can be found in the paragraph containing equation (5), the manuscript surely benefit from a careful review before submitting.

\subsection*{\ovalbox{General Comments}}
I have read the manuscript "EEG characterization and classification based on Histogram of Gradients of Signal Plots" with great interest since its title promised a solution to an important problem. In spite of presenting an interesting methodology and some interesting issues related to EEG characterization, I consider the manuscript is not ready for publication. 
\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for your feedback. We have modified several points according to your recommendations, we hope the article is now better suited for publication. 
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

2.1) the title is misleading since a first, and naive, read suggest being able to classify and characterize any EEG with their different components (i.e., P200, P300, P600, N100, N200, N400 to name a few). Clearly, the title does not match the content since the authors focus on the P300

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation. The title of the article was change accordingly to your suggestion.  
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

2.2) there is evidence of the benefits of working with these signals P300, but, is the proposed method extensible to other signals?, this is not clear from the manuscript.

### 2017-12-04 article.tex
Cambios en la Intro y la carta

BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon  \citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.  New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments \citep{Huggins2016}.  However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement \citep{Clerc}.

In this work, a new method to characterize and classify EEG signals is presented, expanded and detailed.  Its validity is verified by processing offline data for ALS patients.  This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by P300 Event Related Potential \citep{Knuth2006}.

The method is based on the morphological analysis of the shape of the EEG signal \citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots \citep{Hartman2005}.

This paper reports a method to, (1) characterize EEG signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) how this characterization can be used to implement a BCI classification scheme to identify Event Related Potentials, particularly the well-known P300, on an offline and public dataset.

@@

\begin{equation}
\hat{r} = \arg \min_u \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2,\; u = 1,\dots,6
\label{eq:multiclassificationrow}
\end{equation}

@@

\begin{equation}
\hat{c} = \arg \min_u \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2,\; u = 7,\dots,12
\label{eq:multiclassificationcol}
\end{equation}

where $q$ is a template descriptor that belongs to the $k$ nearest neighbors  $NN_T(d^c_u)$.  By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.

>>>>

BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon  \citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.  New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments \citep{Huggins2016}.  However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement~\citep{Clerc}.

In~\citep{Ramele2016}, the authors introduces a method for classification of rhythmic patterns using gradient histograms orientations. Inspired in that work, we proposed another application of the developed method for classify and describe transient events like those produced by P300 Event Related Potential. 
 Its validity is verified by processing offline data for ALS patients.  
The proposed method is based on the morphological analysis of the shape of the EEG signal~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and it was built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots~\citep{Hartman2005}.

This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) describe the way in which  this classification can be used to implement a BCI Event Related Potentials identification, on an offline and public dataset and another own dataset.

@@

\begin{equation}
\hat{r} = \arg \min_{u \in \{1,\dots,6\}} \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2
\label{eq:multiclassificationrow}
\end{equation}

@@

\begin{equation}
\hat{c} = \arg \min_{u \in \{7,\dots,12\}} \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2
\label{eq:multiclassificationcol}
\end{equation}

where $NN_T(d^l_u), \;l=r,c$ is the set of the $k$ nearest neighbors to $d^l_u, \;l=r,c$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l=r,c$.  By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
